---
title: "Exploratory Data Analysis Report"
output:
  html_document: 
    theme: cerulean
    toc: yes
    toc_depth: 6
params:
  data: data
  Target: Target
  label: label
  sc: sc
  sn: sn
  Rc: Rc
---
  
```{r setup, include=FALSE}
library(rmarkdown)
library(SmartEDA)
library(knitr)
library(scales)
library(gridExtra)
library(ggplot2)

data <- params$data

```
  
## Exploratory Data analysis (EDA)
Analyzing the data sets to summarize their main characteristics of variables, often with visual graphs, without using a statistical model. 

### 1. Overview of the data
Understanding the dimensions of the dataset, variable names, overall missing summary and data types of each variables


```{r od_1,warning=FALSE,eval=F,include=T}
# Overiview of the data
ExpData(data=data,type=1)
# Structure of the data
ExpData(data=data,type=2)
```

```{r od_2,warning=FALSE,eval=T,include=F}
ovw_tabl <- ExpData(data=data,type=1)
ovw_tab2 <- ExpData(data=data,type=2)
```

**Overview of the data**
```{r od_3,warning=FALSE,eval=T,render=ovw_tabl,echo=F}
paged_table(ovw_tabl)
```

**Structure of the data**
```{r od_32,warning=FALSE,eval=T,render=ovw_tab2,echo=F}
paged_table(ovw_tab2)
```

**Target variable**

Summary of continuous dependent variable

  1. Variable name - **`r Target `**
  2. Variable description - **`r label `**

```{r dv-r,warning=FALSE,eval=T,include=F}
summary(data[,Target])
```

### 2. Summary of numerical variables

```{r snv_all,warning=FALSE,eval=T,include=F}
snv_2 = ExpNumStat(data,by="GA",gp=Target,Qnt=seq(0,1,0.1),MesofShape=2,Outlier=TRUE,round=2)
rownames(snv_2)<-NULL
```

Summary statistics when dependent variable is Continuous **`r Target `**. 

```{r snv_2, warning=FALSE,eval=F,include=T}
ExpNumStat(data,by="A",gp=Target,Qnt=seq(0,1,0.1),MesofShape=2,Outlier=TRUE,round=2)
```

```{r snv_2.1,warning=FALSE,eval=T,render=snv_2,echo=F}
paged_table(snv_2)
```

### 3. Distributions of numerical variables
Graphical representation of all numeric features, used below types of plots to explore the data

  * Quantile-quantile plot (Univariate)
  * Density plot (Univariate)
  * Scatter plot (Bivariate)

#### Quantile-quantile plot for Numerical variables - Univariate  
Quantile-quantile plot for all Numerical variables

```{r snv2_new,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpOutQQ(data,nlim=4,fname=NULL,Page=c(2,2),sample=sn)
```
    
#### Density plots for numerical variables - Univariate
Density plot for all numerical variables

```{r snv2,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpNumViz(data,gp=NULL,nlim=5,fname=NULL,col=NULL,Page=c(2,2),sample=sn)
```

#### Scatter plot for all Numeric variables vs Continuous dependent variable
Scatter plot between all numeric variables and target variable **`r Target `**.
This plot help to examine how well a target variable is correlated with list of dependent variables in the data set.

Dependent variable is **`r Target `** (continuous). 

```{r snv1,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpNumViz(data,gp=Target,nlim=5,fname=NULL,col=NULL,Page=c(2,2),sample=sn)
```

#### Correlation between Target and Independent variables
```{r eda_400, eval=T,include=F}
snv_22 = ExpNumStat(data,by="GA",gp=Target,Qnt=NULL,MesofShape=2,Outlier=FALSE,round=2,dcast=T,val="cor")
rownames(snv_22)<-NULL

```

```{r eda_400.1, warning=FALSE,eval=F,include=T}
ExpNumStat(data,by="GA",gp=Target,Qnt=NULL,MesofShape=2,Outlier=FALSE,round=2,dcast=T,val="cor")
```

```{r eda_400.2,warning=FALSE,eval=T,render=snv_22,echo=F}
paged_table(snv_22)
```

### 4. Summary of categorical variables
Summary of categorical variables

```{r eda_4, eval=T,include=F}
et1 <- ExpCTable(data,Target=NULL,margin=1,clim=10,nlim=NULL,round=2,bin=NULL,per=T)
rownames(et1)<-NULL
```

```{r eda_41, eval=T,include=F}
et11 <- ExpCTable(data,Target=Target,margin=1,clim=10,nlim=NULL,round=2,bin=4,per=T)
rownames(et11)<-NULL
```
  * frequency for all categorical independent variables

```{r e4.1, warning=FALSE,eval=F,include=T}
ExpCTable(data,margin=1,clim=10,nlim=NULL,round=2,bin=NULL,per=T)
```

```{r e4.1.1,warning=FALSE,eval=T,render=et1,echo=F}
paged_table(et1)
```

 * frequency for all categorical independent variables by descretized **`r Target `**

```{r e4.2, warning=FALSE,eval=F,include=T}
##bin=4, descretized 4 categories based on quantiles
ExpCTable(data,Target=Target,margin=1,clim=10,nlim=NULL,round=2,bin=4,per=T)
```

```{r e4.2.1,warning=FALSE,eval=T,render=et11,echo=F}
paged_table(et11)
```

### 5. Distributions of Categorical variables
Graphical representation of all Categorical variables

  * Bar plot (Univariate)
 
Bar plot with vertical or horizontal bars for all categorical variables

```{r bp1,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpCatViz(data,target=NULL,clim=10,margin=2,Page = c(2,2),sample=sc)
```
