---
title: "Exploratory Data Analysis Report"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: cerulean
    toc: yes
    toc_depth: 6
params:
  data: data
  Target: Target
  label: label
  theme: theme
  sc: sc
  sn: sn
  Rc: Rc
---
  
```{r setup, include=FALSE}
library(rmarkdown)
library(SmartEDA)
library(knitr)
library(scales)
library(gridExtra)
library(ggplot2)

data <- params$data

```
  
## Exploratory Data analysis (EDA)
Analyzing the data sets to summarize their main characteristics of variables, often with visual graphs, without using a statistical model. 

### 1. Overview of the data
Understanding the dimensions of the dataset, variable names, overall missing summary and data types of each variables

```{r od_1,warning=FALSE,eval=F,include=T}
# Overiview of the data
ExpData(data=data,type=1)
# Structure of the data
ExpData(data=data,type=2)
```

```{r od_2,warning=FALSE,eval=T,include=F}
ovw_tabl <- ExpData(data=data,type=1)
ovw_tab2 <- ExpData(data=data,type=2)
```

**Overview of the data**
```{r od_3,warning=FALSE,eval=T,render=ovw_tabl,echo=F}
paged_table(ovw_tabl)
```

**Structure of the data**
```{r od_32,warning=FALSE,eval=T,render=ovw_tab2,echo=F}
paged_table(ovw_tab2)
```

**Target variable**

Summary of categorical dependent variable

  1. Variable name - **`r Target `**
  2. Variable description - **`r label `**

```{r dd,warning=FALSE,eval=T,include=F}
if(is.null(label)) label <- Target
tab_tar <- data.frame(table(data[,Target]))
tab_tar$Descriptions <- label
names(tab_tar) <- c(Target,"Frequency","Descriptions")
rownames(tab_tar)<-NULL
```

```{r dv-r,warning=FALSE,eval=T,render=tab_tar,echo=F}
tab_tar
```

### 2. Summary of numerical variables

Summary of all numerical variables 

```{r snv_all,warning=FALSE,eval=T,include=F}
snv_2 = ExpNumStat(data,by="GA",gp=Target,Qnt=seq(0,1,0.1),MesofShape=2,Outlier=TRUE,round=2)
rownames(snv_2)<-NULL
```

Summary statistics when dependent variable is categorical **`r Target `**. Summary statistics will be splitted into category level

```{r snv_2, warning=FALSE,eval=F,include=T}
ExpNumStat(data,by="GA",gp=Target,Qnt=seq(0,1,0.1),MesofShape=2,Outlier=TRUE,round=2)
```

```{r snv_2.1,warning=FALSE,eval=T,render=snv_2,echo=F}
paged_table(snv_2)
```

### 3. Distributions of Numerical variables

* Box plots for all numerical variables vs categorical dependent variable - Bivariate comparision only with categories

  * Quantile-quantile plot(Univariate)
  * Density plot (Univariate)
  * Box plot (Univariate and Bivariate)
  * Scatter plot (Bivariate)

#### Quantile-quantile plot for Numerical variables - Univariate  
Quantile-quantile plot for all Numerical variables

```{r snv2_new,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpOutQQ(data,nlim=4,fname=NULL,Page=c(2,2),sample=sn)
```
      
#### Density plots for Numerical variables - Univariate
Density plot for all Numerical variables

```{r snv2,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpNumViz(data,target=NULL,type=1,Page=c(2,2),theme=theme,sample=sn)
```

#### Scatter plot for all numeric features - Bivariate analysis

```{r snv4223,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpNumViz(data,target=Target,Page=c(2,1),theme=theme,sample=sn,scatter=TRUE)
```

#### Box plots for all numeric features vs categorical dependent variable - Bivariate comparision only with categories
Boxplot for all the numeric attributes by each category of **`r Target `**

```{r snv4,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpNumViz(data,target=Target,type=2,theme=theme,Page=c(2,2),sample=sn)
```

### 4. Summary of categorical variables
Summary of categorical variable

```{r eda_4, eval=T,include=F}
if(is.null(Rc)) {Rc=as.character(paste0(unique(data[,Target])[1]))}
et1 <- ExpCTable(data,Target=Target,margin=1,clim=10,nlim=5,round=2,bin=NULL,per=F)
rownames(et1)<-NULL

et4 <- ExpCatStat(data,Target=Target,result = "Stat",clim=10,nlim=5,bins=10,Pclass=Rc,plot=FALSE,top=20,Round=2)
rownames(et4)<-NULL

et5 <- ExpCatStat(data,Target=Target,result = "IV",clim=10,nlim=5,bins=10,Pclass=Rc,plot=FALSE,top=20,Round=2)
rownames(et5)<-NULL
```

**Cross tabulation with target variable**

  * Custom tables between all categorical independent variables and traget variable **`r Target `** 

```{r e4.1, warning=FALSE,eval=F,include=T}
ExpCTable(data,Target=Target,margin=1,clim=10,nlim=5,round=2,bin=NULL,per=F)
```

```{r e4.1.1,warning=FALSE,eval=T,render=et1,echo=F}
if(!is.null(dim(et1))){
paged_table(et1)  
}

```

**Information Value**

```{r e4.2, warning=FALSE,eval=F,include=T}
ExpCatStat(data,Target=Target,Label=label,result = "IV",clim=10,nlim=5,Pclass=Rc)

```

```{r e4.21,warning=FALSE,eval=T,render=et5,echo=F}
paged_table(et5)
```

**Statistical test**
```{r e4.3, warning=FALSE,eval=F,include=T}
ExpCatStat(data,Target=Target,Label=label,result = "Stat",clim=10,nlim=5,Pclass=Rc)
```

```{r e4.31,warning=FALSE,eval=T,render=et4,echo=F}
paged_table(et4)
```

**Variable importance based on Information value**

```{r ed3.91,warning=FALSE,eval=T,fig.align='center',fig.height=7,fig.width=7}
varimp <- ExpCatStat(data,Target=Target,result = "Stat",clim=10,nlim=5,Pclass=Rc,bins=10,plot=TRUE,top=30,Round=2)
```

### 5. Distributions of categorical variables
Graphical representation of all categorical variables

  * Bar plot (Univariate)
  * Stacked Bar plot (Bivariate)

**Bar plots for all categorical variables**

  * Bar plot with vertical or horizontal bars for all categorical variables 

```{r bp1,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpCatViz(data,target=NULL,fname=NULL,clim=10,margin=2,theme=theme,Page = c(2,1),sample=sc)
```

  * Stacked bar plot with vertical or horizontal bars for all categorical variables 

```{r bp2,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpCatViz(data,target=Target,fname=NULL,clim=10,margin=2,theme=theme,Page = c(2,1),sample=sc)
```
