---
title: "Exploratory Data Analysis Report"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: cerulean
    toc: yes
    toc_depth: 6
params:
  data: data
  Target: Target
  label: label
  theme: theme
  sc: sc
  sn: sn
  Rc: Rc
---
  
```{r setup, include=FALSE}
library(rmarkdown)
library(SmartEDA)
library(knitr)
library(scales)
library(gridExtra)
library(ggplot2)

data <- params$data

```
  
## Exploratory Data analysis (EDA)
Analyzing the data sets to summarize their main characteristics of variables, often with visual graphs, without using a statistical model. 

### 1. Overview of the data
Understanding the dimensions of the dataset, variable names, overall missing summary and data types of each variables

```{r od_1,warning=FALSE,eval=F,include=T}
# Overiview of the data
ExpData(data=data,type=1)
# Structure of the data
ExpData(data=data,type=2)
```

```{r od_2,warning=FALSE,eval=T,include=F}
ovw_tabl <- ExpData(data=data,type=1)
ovw_tab2 <- ExpData(data=data,type=2)
```

**Overview of the data**
```{r od_3,warning=FALSE,eval=T,render=ovw_tabl,echo=F}
paged_table(ovw_tabl)
```

**Structure of the data**
```{r od_32,warning=FALSE,eval=T,render=ovw_tab2,echo=F}
paged_table(ovw_tab2)
```

### 2. Summary of numerical variables
Summary of all numeric variables 
```{r snv_all,warning=FALSE,eval=T,include=F}
snv_2 = ExpNumStat(data,by="A",gp=NULL,Qnt=seq(0,1,0.1),MesofShape=2,Outlier=TRUE,round=2)
rownames(snv_2)<-NULL
```

```{r snv_2, warning=FALSE,eval=F,include=T}
ExpNumStat(data,by="A",gp=NULL,Qnt=seq(0,1,0.1),MesofShape=2,Outlier=TRUE,round=2)
```

```{r snv_2.1,warning=FALSE,eval=T,render=snv_2,echo=F}
paged_table(snv_2)
```

### 3. Distributions of numerical variables
Graphical representation of all numeric features

  * Quantile-quantile plot(Univariate)
  
```{r snv2_new,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpOutQQ(data,nlim=4,fname=NULL,Page=c(2,2),sample=sn)
```
  
  * Density plot (Univariate)

```{r snv2,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpNumViz(data,target=NULL,type=1,nlim=10,fname=NULL,col=NULL,Page=c(2,2),theme=theme,sample=sn)
```

  * Scatter plot (Bivariate)

```{r snv22_1,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpNumViz(data,Page=c(2,1),sample=sn,theme=theme,scatter=TRUE)
```

### 4. Summary of categorical variables
Summary of categorical variables

```{r eda_4, eval=T,include=F}
et1 <- ExpCTable(data,Target=NULL,margin=1,clim=10,nlim=5,round=2,bin=NULL,per=T)
rownames(et1)<-NULL
```

  * frequency for all categorical independent variables

```{r e4.1, warning=FALSE,eval=F,include=T}
ExpCTable(data,Target=NULL,margin=1,clim=10,nlim=5,round=2,bin=NULL,per=T)
```

```{r e4.1.1,warning=FALSE,eval=T,render=et1,echo=F}
paged_table(et1)
```
>*`NA` is Not Applicable*

### 5. Distributions of categorical variables

  * Bar plots for all categorical variables

#### Bar plots for all categorical variables
Bar plot with vertical or horizontal bars for all categorical variables 

```{r bp1,warning=FALSE,eval=T,include=T,fig.align='center',fig.height=10,fig.width=10}
ExpCatViz(data,target=NULL,fname=NULL,clim=10,margin=2,theme=theme,Page = c(2,2),sample=sc)
```

